---
title: "Defining knitr R chunks in a LaTeX document using the Rcode environment.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rcode}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



```{r setup}
library(mbutils)
```

In a LaTeX document to be compiled using the knitr package, R code
chunks are delimited by the <<>>= and \@ symbols. This is a rather
non-standard way of defining text in a LaTeX document, so it can be
incompatible with other packages.  The mbutils package allows code
chunks to be defined instead as an environment, so the R code "plays
nice" with other LaTeX packages.  Defining code chunks this way is
also useful when using emacs as a text editor, with ESS.

# Motivation

I write nearly everything (papers, code, notes) using emacs
(specifically, Aquamacs Emacs for MacOS: www.aquamacs.org.). I like
emacs for the following reasons:
1.  It's a common editors for coding and LaTeX writing, so I need to
    remember only a single set of keystrokes.
2.  R and emacs are very closely integrated using ESS. I can run an R
    session within an emacs window, copy and paste text back and
    forth, and even run code chunks directly from another buffer.  ESS
    also provides font highlighting for R code.
3. When writing a LaTeX document, I can preview LaTeX math *inline*
    using the preview-latex emacs package (now included as part of AucTeX).
    That is, I can see the typset output directly in the file, and not
    have to keep recompiling a PDF file and viewing it in another window.

Previewing LaTeX output is easy when writing a tex file, with no R
code.  However, R code chunks cause preview-latex to fail.  The reason
is that preview-latex does not treat R code differently than LaTeX
commands. For example, preview-latex would see a \$ in a code chunk
and think it is a delimiter for a LaTeX mathematical expression.  Just
one of these occurrences will mess up the previews (and often font
highlighting) for the entire document.

The solution would be to tell preview-latex to ignore anything between
the <<>>= and the \@.  But there is no good way to do that.  However,
it is possible to configure preview-latex to ignore an
*environment*. Defining new environments is straightforward in LaTeX
with a few lines in the preamble of the document.  But then,
knitr would no longer recognize the text in the environment as an R
code chunk.  This package provides a function that defines an Rcode
environment as a knitr "pattern," so knitr will treat anything inside
that environment as a code chunk.

# The Rcode environment

The Rcode environment is the easy part.  Look at the following Rnw
file that embeds some LaTeX math and R code within a single document.

```
\documentclass{article}
\usepackage{amsmath}

%%%%
\usepackage[textmath,displaymath]{preview}
\newenvironment{Rcode}[1][]{\begin{nopreview}}{\end{nopreview}}
\newcommand{\Sexpr}[1]{}
\PreviewMacro*[{{}}]{\Sexpr}
\PreviewEnvironment{align*}  %% optional
%%%%

\begin{document}

Some code:
\begin{Rcode}[eval=TRUE, echo=TRUE]
A <- list(x=1, y=2)
A$x
\end{Rcode}

Now some math:

\begin{align*}
 1+1&=2\\
 \alpha+\beta&=5
\end{align*}

Closing text.

\end{document}
```

The set of 5 lines near the top of this file do the following:
1.  Load the preview package explicitly, and tell it to preview math
    expressions that are both inline (textmath) and in display style
    (displaymath).  There are other options that are described in the
    preview-latex documentation; those are outside the scope of this
    note.
2. Define a new environment, Rcode, that is, from the LaTeX point of
   view, an alias for the nopreview environment defined by
   preview-latex.  All R code chunks will be written within the Rcode
   environment, as a replacement for <<>>= and \@.  The only
   difference between the Rcode and nopreview environments is that
   Rcode takes a set of optional arguments:  the knitr chunk options.
3.  Define a LaTeX macro to delimit inline code.  Since knitr already
    recognizes \textbackslash Sexpr, that's what we will continue to
    use.
4.  The \textbackslash PreviewMacro* macro tells preview-latex to
    *ignore* the Sexpr macro and all of its arguments and
    contents. This line is necessary so preview-latex does not
    mistakenly treat R code as inline math.
5.  Depending on your application, there may be other definitions and
    arguments you want to pass to preview-latex.  For example, when
    I create beamer slides, I tell preview-latex to ignore the frame
    environment, but to typeset the align* environment.  Otherwise,
    preview-latex will try to typeset math environments within the
    frame twice  See the preview-latex documentation for more
    information.

R code chunks are delimited by the Rcode environment.  Any optional
knitr chunk options are optional arguments for the environment.
However, for a reason I have not yet been able to determine, you need
the [], even if you do not have chunk options.  I'm not sure how to
fix this yet.


# Knitting the document in R

To compile the document, use the same knit commands you use now (I
usually use knit2pdf).

knitr decides what elements of the document are code chunks by
scanning for certain text patterns.  The patterns are usually
determined by the file extension of the input file.  For example, if
knitr sees that it is processing an Rnw file, it knows to look for
<<>>= and \@ for code chunks, \textbackslash Sexpr for inline code,
and so forth. Without further action, knitr will ignore the Rcode
environment, because it does not know that our Rcode environment
contains the code.  Even worse, LaTeX may try to typeset the contents
of the environment, which will likely lead to an error.

The mbutils package provides the information knitr needs to recognize
the Rcode environment as a code chunk.  Before knitting the document,
simply call the function

```{r, eval=FALSE, echo=TRUE}
pat_rcode()
```

That function will override the default patterns knitr uses for *all* file
formats.  knitr will scan for \textbackslash begin\{Rcode\}\dots
\textbackslash end\{Rcode\}, and \textbackslash Sexpr, and will
*ignore* <<>>= and \@.  Therefore, if you then want to knit a  file
with "classic Sweave" syntax, you will need to either set that pattern
with `pat_rnw()`, or run `knit_patterns$restore()` and let knitr
figure out the right pattern from the file extension.  To go back to
the new Rcode syntax, just call `pat_rcode()` again.

To summarize:  Within the Rnw file, define R code chunks as

```
\begin{Rcode}[]
## Code goes here
\end{Rcode}
```

As long as the correct preamble statements are present, preview-latex
will ignore the Rcode environment when generating inline previews.
Before compiling, call `pat_rcode()`. Call `knit_patterns$restore()`
to "clear" any existing patterns. See the knitr documentation for
dealing with other document types.


# Font highlighting in emacs



# Configuring an ess polymode

If you are using emacs, and especially ess, you want to preserve the
ability to switch major modes from LaTeX to R and back again depending
on where you are typing in your document.  ESS has deprecated the
noweb mode in favor of polymode.  This section explains what you need
to do for that.

1.  Ensure the following packages are installed and configured in your
    emacs installation:  ess, polymode, poly-R.

2.  Install poly-latex-R package by coping poly-latex-R.el to your
    emacs packages directory.  poly-latex-R.el is available with this
    R package, either from the inst/emacs directory in the source code,
    or from the \$R\_HOME/library/RlatexUtils/emacs directory in the
    installed package on your system. Install that file using the M-x
    package-install-file function from emacs.

3.  Add the following to your emacs preferences file (probably
    ~/.emacs).  If you are using Aquamacs on MacOS, edit
    ~/Library/Preferences/Aquamacs Emacs/Preferences.el.

```
(require 'ess-site)
(load "ess-autoloads")  ;; need to register Rnw files

(require 'poly-R)
(require 'poly-markdown)
(require 'poly-latex-R)

;; recognize Rnw files as poly-latex+R
(add-to-list 'auto-mode-alist '("\\.[rR]nw$" . poly-latex+R-mode))
(add-to-list 'auto-mode-alist '("\\.[rR][Ll]atex$" . poly-latex+R-mode))
;; this means that "classic" noweb Rnw files will not be automatically opened
;; in poly-noweb+R mode. In that case, set the mode locally.
;;   manual:  M-x poly-noweb+R-mode
;;   file variable:    %% -*- mode: poly-noweb+R -*-
;; Can do the same thing to set the new poly-latex+R mode, if needed.


(add-to-list 'LaTeX-verbatim-environments "Rcode")
(add-to-list 'TeX-file-extensions "RLatex")
(add-to-list 'TeX-file-extensions "rlatex")
```



