---
title: "Using RlatexUtils"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RlatexUtils)
```

RlatexUtils is a specialized package to solve a specific problem.
When editing an Rnw file in emacs, one might want to use the
preview-latex emacs package for inline previews of typeset math. The
preview-latex commands often mistake elements of R code for LaTeX math
expressions.  The most common problem I have found is mistaking a \$
in an R list element name for the \$\dots\$ that define LaTeX math.

It is possible to configure preview-latex, in the document preamble,
to ignore expressions contained in certain LaTeX expressions.  But we
cannot tell preview-latex to ignore text between a <<>>= and a
\@. This package tells knitr to recognize the code chunk defined in
the environment.

To put it another way:  Currently, knitr recognizes code chunks
defined by <<>>= and \@, but preview-latex will try to typeset them,
and fail. We tell knitr to recognize code in an Rcode environment,
which we also tell preview-latex to ignore.

# An example of the problem.

Consider the following Rnw file for LaTeX output.

```
\documentclass{article}
\usepackage{amsmath}
\begin{document}

Some code:
<<>>=
A <- list(x=1, y=2)
A$x
@

Now some math:

\begin{align*}
 1+1&=2\\
 \alpha+\beta&=5
\end{align*}

Closing text.

\end{document}
```

This file can be processed by knitr with no problem, but preview-latex
will mistake the \$ in `A\$x` as the start of a LaTeX math
expression.  Since we do not want code chunks to be typeset, we will
instead load latex-preview explicitly, define the Rcode environment
and Sexpr functions, and tell preview-latex to ignore both of them.

```
\documentclass{article}
\usepackage{amsmath}

\usepackage[textmath,displaymath,showlabels]{preview}
\newenvironment{Rcode}[1][]{\begin{nopreview}}{\end{nopreview}}
\newcommand{\Sexpr}[1]{}

\PreviewEnvironment{align*}
\PreviewMacro*[{{}}]{\Sexpr}

\begin{document}

Some code:
\begin{Rcode}[]
A <- list(x=1, y=2)
A$x
\end{Rcode}

Now some math:

\begin{align*}
 1+1&=2\\
 \alpha+\beta&=5
\end{align*}

Closing text.

\end{document}
```
As far as preview-latex is concerned, the Rcode environment is
identical to a nopreview environment, whose contents are skipped
entirely. Further, `\\Sexpr` is just a macro that will also be
ignored. Our inline previews of the LaTeX material should now work as
expected.

But what about knitr, which does not recognize the contents of Rcode
as an R code chunk that should be processed?  That's where this
package comes in.

# Using the package




# Configuring an ess polymode

If you are using emacs, and especially ess, you want to preserve the
ability to switch major modes from LaTeX to R and back again depending
on where you are typing in your document.  ESS has deprecated the
noweb mode in favor of polymode.  This section explains what you need
to do for that.

1.  Ensure the following packages are installed and configured in your
    emacs installation:  ess, polymode, poly-R.

2.  Install poly-latex-R package by coping poly-latex-R.el to your
    emacs packages directory.  poly-latex-R.el is available with this
    R package, either from the inst/emacs directory in the source code,
    or from the \$R\_HOME/library/RlatexUtils/emacs directory in the
    installed package on your system. Install that file using the M-x
    package-install-file function from emacs.

3.  Add the following to your emacs preferences file (probably
    ~/.emacs).  If you are using Aquamacs on MacOS, edit
    ~/Library/Preferences/Aquamacs Emacs/Preferences.el.

```
(require 'ess-site)
(load "ess-autoloads")  ;; need to register Rnw files

(require 'poly-R)
(require 'poly-markdown)
(require 'poly-latex-R)

;; recognize Rnw files as poly-latex+R
(add-to-list 'auto-mode-alist '("\\.[rR]nw$" . poly-latex+R-mode))
(add-to-list 'auto-mode-alist '("\\.[rR][Ll]atex$" . poly-latex+R-mode))
;; this means that "classic" noweb Rnw files will not be automatically opened
;; in poly-noweb+R mode. In that case, set the mode locally.
;;   manual:  M-x poly-noweb+R-mode
;;   file variable:    %% -*- mode: poly-noweb+R -*-
;; Can do the same thing to set the new poly-latex+R mode, if needed.


(add-to-list 'LaTeX-verbatim-environments "Rcode")
(add-to-list 'TeX-file-extensions "RLatex")
(add-to-list 'TeX-file-extensions "rlatex")
```



